\Title

### 2022-11-30 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.
methods

and from the ANR project % N° ANR-22-XXX-XXX 

@@

Let us now formally define the heterogeneous model, as well as the task and how we will learn to solve it. First, we will define the model for the efficient detection of event-based motifs using a feed-forward layer of spiking neurons with heterogeneous delays. Our SNN model will be tested on a computationally relevant task of motion detection in an event-based sensory input. By analogy with biological conditions, we will design a paradigm simulating eye movements where an input image is translated by a parameterized saccadic trajectory. Solving this task is, for example, useful in real-life situations to compensate for eye, head or body movements and to provide robust image categorization by stabilizing the image on the retina. The task is thus to guess the motion as accurately as possible (\seeFig{motion_task}) and we will describe a semi-supervised learning mechanism for weights and delays. To this end, we will extend the SNN model to be efficient for the resolution of the task by including spatio-temporal convolutional kernels. We will finally describe its implementation on conventional computers. 

@@

    \includegraphics[width=0.490\linewidth]{figures/figure_1a_k.pdf}
    \includegraphics[width=0.490\linewidth]{figures/figure_1c.pdf}
    \includegraphics[width=0.490\linewidth]{figures/figure_1b.pdf}
    \includegraphics[width=0.490\linewidth]{figures/figure_1a.pdf}

@@

	    {\bf (b)}~The generative model is defined as a set of motifs (here $4$ PGs) each defined by a set of different synapses defined by weights (increasing with the radius of the black dots) at each different delay. 
	    {\bf (c)}~We will define a layer of neurons that will detect these PGs such that the addresses are the synapses of these neurons (which we therefore call here synapses).

@@

	    {\bf (c)}~The output of the model provides with the predicted probability of occurrence of each PG motif at any time, which may be used to generate a spike as a Bernoulli trial, providing in this particular case with an exact identification of PGs occurrences.
	    {\bf (d)}~Knowing the results of this detection, one may for illustration purposes highlight them by different colors in the raster plots, showing that in this synthetic examples, all spikes are now associated with a PG. 

@@

%\subsubsection{Implementation using spiking neurons with heterogeneous delays}
%
The activation function of our spiking neural is a softmax function implementing a form of  Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022}, in analogy to a spiking Winner-Take-All network~\citep{nessler_bayesian_2013}. 
%It transforms this list of weights into a probability with the following formula:
%$
%Pr(k=\postsynaddr \; \vert \; \timev) =
%\frac 1 Z
%{\exp  (\mathcal{C}^\postsynaddr(\timev) +\bias^\postsynaddr) }
%$ 
%where $\mathcal{C}^\postsynaddr(\timev) = \sum
%\activeweights^\postsynaddr(t)
%$ is the sum of the synaptic weights and $\bias^\postsynaddr$ is the bias linked to neuron $\postsynaddr$. 
%In particular, we expect that some specific motifs may become tightly synchronized as they reach the basal dendritic tree, leading to a high postsynaptic activity which makes it progressively more likely to generate an output spike.
%%
%
\note{say that it is effortless in biology but difficult on conventional computers}
In our MLR model with $\Nclass=\Nspeed$ classes, a probability value is predicted for each event at address $\presynaddr_\arank$ and at time $\timev_\arank$ as a softmax function of the linear combination of the list of events on the basal dendrite of a neuron $\postsynaddr$ in association to a specific class. The linear combination can be defined by a set of synapses $\synapse^\postsynaddr$ as described in the heterogeneous delays model. 
%
\subsubsection{Performance on synthetic data}
%
\begin{figure}%[t!]
    \centering
    \includegraphics[width=0.490\linewidth]{figures/figure_N_PGs.pdf}
    \includegraphics[width=0.490\linewidth]{figures/figure_N_PG_time.pdf}
	    \caption{Detecting event-based motifs using spiking neurons with heterogeneous delays. 
	    {\bf (a)}~Accuracy of PG detection as a function of the number $M$ of kernels.
	    {\bf (b)}~Accuracy of PG detection as a function of the temporal depth $D$ of kernels among $M=144$ kernels.
	    	     }
    \label{fig:model_results}
\end{figure}
% 
This generative model defined above allows to determine this inference model for guessing sources $B$ when observing a raster plot $A$. This assumes that we know the PGs as defined by the $W_b$ matrices. The underlying metric is the binary cross-entropy, as used in the logistic regression model. In particular, if we consider PG kernels with similar decreasing exponential time profile, we can prove that this is similar to the method of~\citet{berens_fast_2012}. In our specific case, the difference is that the regression is performed in both dendritic and delay space by extending the summation using a temporal convolution operator. Using this forward model, it is possible to estimate the logit (inverse of a sigmoid) $\hat{B}(b, t)$ for the presence of a PG of address $b$ and at time $t$ by using the transpose convolution operator. It thus comes that when observing $A$, then one may infer $\hat{B} = A \ast W^T$ and select the most activated items.

>>>>

and from A*Midex grant number AMX-21-RID-025

@@

Let us now formally define the spiking model with  heterogeneous delays, as well as the task and how we will learn to solve it. First, we will define the model for the efficient detection of event-based motifs using a feed-forward layer of spiking neurons with heterogeneous delays. Our SNN model will be tested on a computationally relevant task of motion detection in an event-based sensory input. By analogy with biological conditions, we will design a paradigm simulating eye movements where an input image is translated by a parameterized saccadic trajectory. Solving this task is, for example, useful in real-life situations to compensate for eye, head or body movements and to provide robust image categorization by stabilizing the image on the retina. The task is thus to guess the motion as accurately as possible (\seeFig{motion_task}) and we will describe a semi-supervised learning mechanism for weights and delays. To this end, we will extend the SNN model to be efficient for the resolution of the task by including spatio-temporal convolutional kernels. We will finally describe its implementation on conventional computers. 

@@

    \includegraphics[width=0.490\linewidth]{figures/THC_1a_k.pdf}
    \includegraphics[width=0.490\linewidth]{figures/THC_1c.pdf}
    \includegraphics[width=0.490\linewidth]{figures/THC_1b.pdf}
    \includegraphics[width=0.490\linewidth]{figures/THC_1a.pdf}

@@

	    {\bf (b)}~We show these motifs, each identified at the top by a different color. Each neurons and $71$ different possible delays is assigned the evidence of activation (red) or deactivation (blue). %The generative model is defined as a set of motifs (here $4$ PGs) each defined by a set of different synapses defined by weights (increasing with the radius of the black dots) at each different delay. 
	    {\bf (c)}~The activation in time of the different motifs is then used to define a generative model for drawing a raster plot on the multi-unit address space.
        % We will define a layer of neurons that will detect these PGs such that the addresses are the synapses of these neurons (which we therefore call here synapses).

@@

	    % {\bf (c)}~The output of the model provides with the predicted probability of occurrence of each PG motif at any time, which may be used to generate a spike as a Bernoulli trial, providing in this particular case with an exact identification of PGs occurrences.
	    {\bf (d)}~Knowing the results of this detection, one may for illustration purposes highlight them by different colors in the raster plots, showing that in this synthetic examples, all spikes are now associated with a PG.  The original raster plot can be annotated with each identified spiking motif. 

@@

%\subsubsection{Implementation using spiking neurons with heterogeneous delays}
%
The activation function of our spiking neural is a softmax function implementing a form of  Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022}, in analogy to a spiking Winner-Take-All network~\citep{nessler_bayesian_2013}. 
%It transforms this list of weights into a probability with the following formula:
%$
%Pr(k=\postsynaddr \; \vert \; \timev) =
%\frac 1 Z
%{\exp  (\mathcal{C}^\postsynaddr(\timev) +\bias^\postsynaddr) }
%$ 
%where $\mathcal{C}^\postsynaddr(\timev) = \sum
%\activeweights^\postsynaddr(t)
%$ is the sum of the synaptic weights and $\bias^\postsynaddr$ is the bias linked to neuron $\postsynaddr$. 
%In particular, we expect that some specific motifs may become tightly synchronized as they reach the basal dendritic tree, leading to a high postsynaptic activity which makes it progressively more likely to generate an output spike.
%%
%
\note{say that it is effortless in biology but difficult on conventional computers}
In our MLR model with $\Nclass=\Nspeed$ classes, a probability value is predicted for each event at address $\presynaddr_\arank$ and at time $\timev_\arank$ as a softmax function of the linear combination of the list of events on the basal dendrite of a neuron $\postsynaddr$ in association to a specific class. The linear combination can be defined by a set of synapses $\synapse^\postsynaddr$ as described in the heterogeneous delays model. 
%
\subsubsection{Performance on synthetic data}
%
\begin{figure}%[t!]
    \centering
    \includegraphics[width=0.490\linewidth]{figures/THC_N_PGs.pdf}
    \includegraphics[width=0.490\linewidth]{figures/THC_N_PG_time.pdf}
	    \caption{Detecting event-based motifs using spiking neurons with heterogeneous delays. 
	    {\bf (a)}~Accuracy of PG detection as a function of the number $M$ of kernels.
	    {\bf (b)}~Accuracy of PG detection as a function of the temporal depth $D$ of kernels among $M=144$ kernels.
	    	     }
    \label{fig:model_results}
\end{figure}
% 
This generative model defined above allows to determine this inference model for guessing sources $B$ when observing a raster plot $A$. This assumes that we know the PGs as defined by the $W_b$ matrices. The underlying metric is the binary cross-entropy, as used in the logistic regression model. In particular, if we consider PG kernels with similar decreasing exponential time profile, we can prove that this is similar to the method of~\citet{berens_fast_2012}. In our specific case, the difference is that the regression is performed in both dendritic and delay space by extending the summation using a temporal convolution operator. Using this forward model, it is possible to estimate the logit (inverse of a sigmoid) $\hat{B}(b, t)$ for the presence of a PG of address $b$ and at time $t$ by using the transpose convolution operator. It thus comes that when observing $A$, then one may infer $\hat{B} = A \ast W^T$ and select the most activated items.
